---
title: "Descriptive statistics of Paridae monitoring data"
author: "Freya Coursey"
format: html
editor: visual
echo: false
---

## Load packages

```{r}
#| label: Load Packages
#| include: false
# Core packages for data manipulation, visualization, and modeling
library(tidyverse)       # Data manipulation
library(lme4)        # Mixed-effects models
library(broom.mixed) # Tidy methods for mixed models
library(TMB)         # Template Model Builder
library(MuMIn)       # Model selection and multimodel inference
library(gt)          # Table visualisation

# Visualization and theme extensions
library(ggfortify)   # Auto-plotting for statistical results
library(ggiraphExtra) # Interactive ggplot extensions
library(ggthemes)    # Additional themes for ggplot2

# Model interpretation and effects
library(effects)     # Visualizing model effects
library(car)         # Companion to Applied Regression
library(jtools)      # Tools for regression models

# Additional data handling and formatting
library(Hmisc)       # Miscellaneous data utilities
library(pander)      # Rendering tables in markdown
library(conflicted)

conflict_prefer_all("dplyr", quiet = TRUE) # use dplyr if there is a name conflict

# default plot theme
theme_set(theme_classic())
```

## Load data

```{r}
#| label: Load data
#| include: false
#| 
Main_data_milde <- read_csv("Main_data_milde.csv")
View(Main_data_milde)
```

## Data Analysis

Data were analysed in R version r paste(R.version$major, R.version$minor, sep = ".") [@R], using tidyverse for data manipulation and plots [@tidyverse].

```{r}
#| label: Clean and filter data
#| include: false

# File name for cleaning
Nestbox_data <- Main_data_milde

# Covariate to factor
Nestbox_data$factor_Year<-factor(Nestbox_data$Year) # creating a new column
Nestbox_data$Nestbox<-factor(Nestbox_data$Nestbox)
Nestbox_data$Species<-factor(Nestbox_data$Species)

# Factor to covariate
Nestbox_data$LayDate <- as.numeric(as.character(Nestbox_data$LayDate))
Nestbox_data$BroodSize <- as.numeric(as.character(Nestbox_data$BroodSize))
Nestbox_data$Fledged <- as.numeric(as.character(Nestbox_data$Fledged))
Nestbox_data$ClutchSize <- as.numeric(as.character(Nestbox_data$ClutchSize))
Nestbox_data$Hatch_gap <- as.numeric(as.character(Nestbox_data$Hatch_gap))

# Subsetting data by species
BT.data<-subset(Nestbox_data, Species=="Blåmeis")
GT.data<-subset(Nestbox_data, Species=="Kjøttmeis")

# Subsetting by occupancy
occupied.data<-subset(Nestbox_data, Occupied=="yes")
occupied.data <- filter(occupied.data, LayDate!= ".")
occupied.data <- filter(occupied.data, Species!= ".")
occupied.data <- filter(occupied.data, Species!= "?")

```

## Descriptive statistics

In total, we monitored `r nrow(occupied.data)` nests of birds in nestboxes from `r min(Main_data_milde$Year)` to `r max(Main_data_milde$Year)`.

We monitored `r sum(occupied.data$Species =="Blåmeis")` nests of blue tit, and `r sum(occupied.data$Species =="Kjøttmeis")` nests of great tit. In 2022, one nest had both blue and great tit young raised by blue tit parents. The remaining rests were of coal tit and pied flycatcher (@tbl-species-laydate). 

```{r}
#| label: occupancy-rate-table

#Occupancy rate
group_counts <- table(Nestbox_data$Occupied)
total_count <- sum(group_counts)
percentage_by_group <- (group_counts / total_count) * 100
# Combine group names and percentages into a data frame
occupancy_rate_table <- data.frame(
  Number = group_counts,
  Occupied = names(percentage_by_group),
  Percentage = percentage_by_group)
print(occupancy_rate_table)
```

```{r}
#| label: species-percentages-table

#Species percentage
group_counts <- table(occupied.data$Species)
total_count <- sum(group_counts)
percentage_by_group <- (group_counts / total_count) * 100
# Combine group names and percentages into a data frame
species_number_table <- data.frame(
  Number = group_counts,
  Species = names(percentage_by_group),
  Percentage = percentage_by_group)
print(species_number_table)
```


```{r}
#| label: tbl-species-laydate
#| tbl-cap: Mean lay date for each species

occupied.data |> 
  group_by(Species) |> 
  summarise(nests = n(), 
            success = sum(Fledged > 0, na.rm = TRUE),
            success_rate = round((success/nests)*100, digits=0),
            mean_laydate = round(mean(LayDate), digits=0),
            mean_std_dev = round(sd(LayDate), digits = 2), 
            .groups = "drop") |> 
  mutate(mean_laydate = paste(mean_laydate, "(+/-)", mean_std_dev)) |> 
  gt() |> 
  cols_label(nests = "Number of nests", success = "Number of successful nests", success_rate = "Success rate %", Species = "Species", mean_laydate = "Mean Lay Date", mean_std_dev = "Standard deviation")
  
```

```{r}
#| label: fig-species-laydate
#| fig-alt: "A box plot of species lay dates"
#| fig-width: 5.2

# exclude kjøttmeis/blåmeis
ex.occupied.data <- filter(occupied.data, Species!= "Kjøttmeis/Blåmeis")
ggplot(ex.occupied.data, aes(x = Species, y = LayDate)) +
  geom_boxplot(fill = "lightgreen") +
  labs(x = "Species", y = "Lay Date (day since 1st April)")
```

